Gammapy: A Python package for gamma-ray astronomy

### 2023-06-13 src/ms-review.tex
Reduce repetition in data reduction description

After all data have been projected onto the same geometry, it is typically
required to improve the residual hadronic background estimate. As residual hadronic
background models can be subject to significant systematic uncertainties,
these models can be improved by taking into account actual data
from regions without known \gammaray sources. This includes methods 
such as the ring or the FoV background techniques or background measurements
performed within, e.g. reflected regions~\citep{Berge07}.

@@

Observations can be either modeled individually, or in a joint likelihood analysis.
In the latter case the total (joint) log-likelihood is given by the sum of the log-likelihoods
per observation $\mathcal{L} = \sum{i=0}^M \mathcal{C_i}$ for M observations. Other fit
statistics such as \textit{WStat}, where the background is estimated from an independent
measurement or a simple $\chi^2$ for flux points are also often used
and are supported by \gammapy as well. Models can be simple analytical models or more complex
ones from radiation mechanisms of accelerated particle populations (e.g. inverse Compton or $\pi^{0}$ decay).

@@

The DL3 level data represented by the \code{Observation} class consist
of two types of elements: first, a list of \gammaray events with relevant physical
quantities such as estimated energy, direction and arrival
times, which is represented by the \code{EventList} class. Second, a set of
associated IRFs, providing the response of the system, typically
factorised in independent components as described in
Section~\ref{ssec:gammapy-irf}. The separate handling of event lists and IRFs
additionally allows for data from non-IACT \gammaray instruments to be read. For
example, to read \fermi data, the user can read separately their event list
(already compliant with the \gadf specifications) and then find the appropriate
IRF classes representing the response functions provided by \fermi, see
example in Section~\ref{ssec:multi-instrument-analysis}.

@@

\subsubsection{Energy Dispersion}
For \iacts, the energy resolution and bias, sometimes called energy dispersion,
is typically parametrised in terms of the so-called
migration parameter ($\mu$, see section \label{sec:gammaray-data-analysis}).
 However, more complex
shapes are also common. The distribution of the migration parameter is given at each offset angle and
true energy. The main sub-classes are the \code{EnergyDispersion2D} which is
designed to handle the raw instrument description, and the \code{EDispKernelMap},
which contains an energy dispersion matrix per sky position, i.e., a 4-dimensional
sky map where each position is associated to an energy dispersion matrix.
The energy dispersion matrix is a representation of the energy resolution
as a function of the true energy only and implemented in \gammapy
by the sub-class \code{EDispKernel}.

@@

The predicted counts are computed by convolution of the models with the associated
IRFs. Fitting of precomputed flux points is enabled through \code{FluxPointsDataset},
using \emph{$\chi^2$} statistics. Multiple datasets of same or different types can be
bundled together in \code{Datasets} (e.g., Figure \ref{fig*:minted:gp_datasets}),
where the likelihood from each constituent member is added, thus facilitating
joint fitting across different observations, and even different instruments
across different wavelengths. Datasets also provide functionalities for
manipulating reduced data, e.g. stacking, sub-grouping, plotting. Users can
also create their customized datasets for implementing modified likelihood
methods.

@@

\subsection{gammapy.makers}
\label{ssec:gammapy-makers}
%
\begin{figure}
	\small
    \import{code-examples/generated/}{gp_makers}
	\caption{
        Using \code{gammapy.makers} to reduce DL3 level data into a
		\code{MapDataset}. All \code{Maker} classes represent 
		a step in the data reduction process. They take
        the configuration on initialisation of the class. They 
		also consistently define \code{.run()} methods, which take
		a dataset object and optionally an \code{Observation} 
		object. In this way, \code{Maker} classes can be chained
		to define more complex data reduction pipelines. The output
		of the code example is shown in Figure~\ref{fig:code_example_gp_makers}.
    }
	\label{fig*:minted:gp_makers}
\end{figure}
%
The \code{gammapy.makers} sub-package contains the various classes and functions required
to process and prepare \gammaray data from the DL3 to the DL4, representing the input for modelling and fitting.
First, events are binned and IRFs are interpolated and projected onto the chosen analysis
geometry. The end product of the data reduction process is a set of binned counts,
background exposure, psf and energy dispersion maps at the DL4 level. 
The \code{MapDatasetMaker} and \code{SpectrumDatasetMaker} are
responsible for this task for 3D and 1D analyses, respectively (see Figure~\ref{fig*:minted:gp_makers}).

Because background models usually suffer from strong uncertainties, it is required
to correct them from the data themselves. Several techniques are commonly used
in TeV \gammaray astronomy such as FoV background normalization or
background measurement in reflected regions, see~\cite{Berge07}.
Specific \code{Makers} such as the \code{FoVBackgroundMaker} or the
\code{ReflectedRegionsBackgroundMaker} are in charge of this process.
%% CN: I would add a single line explaining what each of them do

>>>>

After all data have been projected onto the same geometry, it is typically
required to improve the residual hadronic background estimate. As residual hadronic
background models can be subject to significant systematic uncertainties,
these models can be improved by taking into account actual data
from regions without known \gammaray sources. This includes methods 
such as the ring or the FoV background normalisation techniques or background measurements
performed within, e.g. reflected regions~\citep{Berge07}.

@@

Observations can be either modeled individually, or in a joint likelihood analysis.
In the latter case the total (joint) log-likelihood is given by the sum of the log-likelihoods
per observation $\mathcal{L} = \sum{i=0}^M \mathcal{C_i}$ for M observations. Other fit
statistics such as \emph{WStat}~\citep{WStat}, where the background is estimated from an independent
measurement or a simple $\chi^2$ for flux points are also often used
and are supported by \gammapy as well. Models can be simple analytical models or more complex
ones from radiation mechanisms of accelerated particle populations (e.g. inverse Compton or $\pi^{0}$ decay).

@@

The DL3 level data represented by the \code{Observation} class consist
of two types of elements: first, the list of \gammaray events 
which is represented by the \code{EventList} class. Second, a set of
associated IRFs, providing the response of the system, typically
factorised in independent components as described in
Section~\ref{ssec:gammapy-irf}. The separate handling of event lists and IRFs
additionally allows for data from non-IACT \gammaray instruments to be read. For
example, to read \fermi data, the user can read separately their event list
(already compliant with the \gadf specifications) and then find the appropriate
IRF classes representing the response functions provided by \fermi, see
example in Section~\ref{ssec:multi-instrument-analysis}.

@@

\subsubsection{Energy Dispersion}
For \iacts, the energy resolution and bias, sometimes called energy dispersion,
is typically parametrised in terms of the so-called
migration parameter ($\mu$, see section \label{sec:gammaray-data-analysis}).
The distribution of the migration parameter is given at each offset angle and
true energy. The main sub-classes are the \code{EnergyDispersion2D} which is
designed to handle the raw instrument description, and the \code{EDispKernelMap},
which contains an energy dispersion matrix per sky position, i.e., a 4-dimensional
sky map where each position is associated to an energy dispersion matrix.
The energy dispersion matrix is a representation of the energy resolution
as a function of the true energy only and implemented in \gammapy
by the sub-class \code{EDispKernel}.

@@

Fitting of precomputed flux points is enabled through \code{FluxPointsDataset},
using \emph{$\chi^2$} statistics. Multiple datasets of same or different types can be
bundled together in \code{Datasets} (e.g., Figure \ref{fig*:minted:gp_datasets}),
where the likelihood from each constituent member is added, thus facilitating
joint fitting across different observations, and even different instruments
across different wavelengths. Datasets also provide functionalities for
manipulating reduced data, e.g. stacking, sub-grouping, plotting. Users can
also create their customized datasets for implementing modified likelihood
methods.

@@

\subsection{gammapy.makers}
\label{ssec:gammapy-makers}
%
\begin{figure}
	\small
    \import{code-examples/generated/}{gp_makers}
	\caption{
        Using \code{gammapy.makers} to reduce DL3 level data into a
		\code{MapDataset}. All \code{Maker} classes represent 
		a step in the data reduction process. They take
        the configuration on initialisation of the class. They 
		also consistently define \code{.run()} methods, which take
		a dataset object and optionally an \code{Observation} 
		object. In this way, \code{Maker} classes can be chained
		to define more complex data reduction pipelines. The output
		of the code example is shown in Figure~\ref{fig:code_example_gp_makers}.
    }
	\label{fig*:minted:gp_makers}
\end{figure}
%
The \code{gammapy.makers} sub-package contains the various classes and functions required
to process and prepare \gammaray data from the DL3 to the DL4.
The end product of the data reduction process is a set of binned counts,
background exposure, psf and energy dispersion maps at the DL4 level, bundled
into a \code{MapDataset} object.
The \code{MapDatasetMaker} and \code{SpectrumDatasetMaker} are
responsible for this task for 3D and 1D analyses, respectively (see Figure~\ref{fig*:minted:gp_makers}).

The correction of background models from the data themselves is supported 
by specific \code{Maker} classes such as the \code{FoVBackgroundMaker} or the
\code{ReflectedRegionsBackgroundMaker}. The former is used to estimate the
normalisation of the background model from the data themselves, while the
latter is used to estimate teh background from regions reflected from the
poiting position.
%% CN: I would add a single line explaining what each of them do
